---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# HoardR

<!-- badges: start -->
[![R-CMD-check](https://github.com/basilinb/HoardR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/basilinb/HoardR/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

HoardR is a unified toolkit for **organising project directories, caching R
objects, and saving tables and figures**. It provides a consistent folder
structure for bioinformatics projects, smart caching that loads from disk if
available, and helpers for saving figures and flat tables.

Key features:

- **Project & experiment directories** via `init_project_dirs()` and
  `init_experiment_dirs()`
- **RDS** for complex objects (DGEList, EList, model fits, lists)
- **CSV/TSV** for flat tables (data.frames, tibbles)
- **PDF/PNG/SVG** and more for figures via ggplot2
- **Consistent file naming** with `make_path()` and `safe_filename()`
- **Skip existing files** to avoid redundant computation

## Installation

You can install the development version of HoardR from
[GitHub](https://github.com/) with:
``` r
# install.packages("pak")
pak::pak("basilinb/HoardR")
```

## Core Functions

### Directory initialisation

| Function                  | Description                                              |
|---------------------------|----------------------------------------------------------|
| `init_project_dirs()`     | Create standard project folder layout under Box storage  |
| `init_experiment_dirs()`  | Create experiment-specific plot and result subdirectories|

### Caching and saving

| Function               | Description                                          |
|------------------------|------------------------------------------------------|
| `make_path()`          | Build consistent paths: `dir/project_analysis.ext`   |
| `safe_filename()`      | Build collision-safe file paths (used by `make_path()`)|
| `cache()`              | Load if cached, otherwise compute + save             |
| `save_figures()`       | Save ggplot/patchwork to one or more files           |
| `cache_with_figures()` | Cache data + save figures in one call                |

## Examples

### Set up a project
```{r example-project-dirs, eval = FALSE}
library(HoardR)

# Create standard folder layout under Box
proj <- init_project_dirs(
  project_number = "P513-5",
  box_base = "~/Library/CloudStorage/Box-Box/BRI_Projects"
)
# proj$root, proj$plots, proj$results, proj$data_clean, proj$data_raw

# Create experiment-specific subdirectories
exp_dirs <- init_experiment_dirs(
  proj,
  experiment_name = "IL13_dose",
  plot_subdirs    = c("volcano", "heatmap", "enrichment",
                      "scatterplot", "boxplot", "GAM"),
  result_subdirs  = c("GSEA", "Norm")
)
# exp_dirs$plots_volcano, exp_dirs$results_GSEA, etc.
```

### Build paths with `make_path()`
```{r example-make-path, eval = FALSE}
# Single path
make_path("P01", "voom_clean", "rds", "DataClean")
# -> "DataClean/P01_voom_clean.rds"

# Multiple extensions at once (returns named vector)
make_path("P01", "results", c("csv", "rds"), "Results")
# -> c(csv = "Results/P01_results.csv", rds = "Results/P01_results.rds")

# safe_filename() is called internally by make_path() to avoid
# collisions when a file with that name already exists
```

### Cache complex objects
```{r example-cache, eval = FALSE}
# Cache a DGEList, EList, or any complex object
dat_clean <- cache(
  make_path(project, "voom_clean", "rds", dir_data),
  {
    # This block only runs on cache miss
    dat <- process_raw_data(dat_raw)
    dat
  }
)
```

### Cache flat tables to CSV + RDS
```{r example-cache-table, eval = FALSE}
# Save to both CSV (for sharing) and RDS (for R)
results <- cache(
  make_path(project, "results", c("csv", "rds"), dir_results),
  {
    run_analysis(dat_clean)
  }
)
```

### Save figures
```{r example-figures, eval = FALSE}
# Save to multiple formats; skips if files already exist
save_figures(
  p_pca,
  paths  = make_path(project, "pca", c("pdf", "png"), dir_figs),
  width  = 8,
  height = 6
)
```

### Cache data + figures together
```{r example-cache-figures, eval = FALSE}
# One compute block -> cache data + save figures
dat_clean <- cache_with_figures(
  data_paths   = make_path(project, "voom_clean", "rds", dir_data),
  figure_paths = make_path(project, "pca", c("pdf", "png"), dir_figs),
  width  = 8,
  height = 6,
  {
    obj <- build_voom(dat_raw)
    list(data = obj, plot = plot_pca(obj))
  }
)
```

## Design Principles

1. **Consistent structure** — `init_project_dirs()` and `init_experiment_dirs()`
   enforce a shared folder layout across all projects and experiments.
2. **Complex objects require RDS** — Attempting to save a DGEList, EList, or
   nested list to CSV/TSV errors immediately with a clear message.
3. **Figures are cached too** — If the output file exists and
   `overwrite = FALSE`, `ggsave()` is skipped entirely.
4. **Flexible compute blocks** — Pass either a bare `{ }` expression or a
   `function()` / `\()` lambda.
5. **Consistent naming** — Always use
   `make_path(project, analysis, ext, dir)` to build paths.

## License

MIT © Basilin Benson