---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# HoardR

<!-- badges: start -->
<!-- badges: end -->

HoardR is a unified toolkit for **caching and saving R objects, tables, and figures**. It provides smart caching that loads from disk if available, otherwise computes and saves. 

Key features:

- **RDS** for complex objects (DGEList, EList, model fits, lists)
- **CSV/TSV** for flat tables (data.frames, tibbles)
- **PDF/PNG/SVG** and more for figures via ggplot2
- **Consistent file naming** with `make_path()`
- **Skip existing files** to avoid redundant computation

## Installation

You can install the development version of HoardR from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("basilinb/HoardR")
```

## Core Functions

| Function | Description |
|----------|-------------|
| `make_path()` | Build consistent paths: `dir/project_analysis.ext` |
| `cache()` | Load if cached, otherwise compute + save |
| `save_figures()` | Save ggplot/patchwork to one or more files |
| `cache_with_figures()` | Cache data + save figures in one call |
| `safe_filename()` | Build collision-safe file paths |

## Examples

### Build paths with `make_path()`

```{r example-make-path, eval = FALSE}
library(HoardR)

# Single path
make_path("P01", "voom_clean", "rds", "DataClean")
# → "DataClean/P01_voom_clean.rds"

# Multiple extensions at once (returns named vector)
make_path("P01", "results", c("csv", "rds"), "Results")
# → c(csv = "Results/P01_results.csv", rds = "Results/P01_results.rds")
```

### Cache complex objects

```{r example-cache, eval = FALSE}
# Cache a DGEList, EList, or any complex object
dat_clean <- cache(
  make_path(project, "voom_clean", "rds", dir_data),
  {
    # This block only runs on cache miss
    dat <- process_raw_data(dat_raw)
    dat
  }
)
```

### Cache flat tables to CSV + RDS

```{r example-cache-table, eval = FALSE}
# Save to both CSV (for sharing) and RDS (for R)
results <- cache(
  make_path(project, "results", c("csv", "rds"), dir_results),
  {
    run_analysis(dat_clean)
  }
)
```

### Save figures

```{r example-figures, eval = FALSE}
# Save to multiple formats; skips if files already exist
save_figures(
  p_pca,
  paths = make_path(project, "pca", c("pdf", "png"), dir_figs),
  width = 8, height = 6
)
```

### Cache data + figures together

```{r example-cache-figures, eval = FALSE}
# One compute block → cache data + save figures
dat_clean <- cache_with_figures(
  data_paths   = make_path(project, "voom_clean", "rds", dir_data),
  figure_paths = make_path(project, "pca", c("pdf", "png"), dir_figs),
  width = 8, height = 6,
  {
    obj <- build_voom(dat_raw)
    list(data = obj, plot = plot_pca(obj))
  }
)
```

## Design Principles

1. **Complex objects require RDS** — Attempting to save a DGEList, EList, or nested list to CSV/TSV errors immediately with a clear message.

2. **Figures are cached too** — If the output file exists and `overwrite = FALSE`, `ggsave()` is skipped entirely.

3. **Flexible compute blocks** — Pass either a bare `{ }` expression or a `function()` / `\()` lambda.

4. **Consistent naming** — Always use `make_path(project, analysis, ext, dir)` to build paths.

## License

MIT © Basilin Benson
